Perturbing SU($n$)-symmetric interactions

### 2020-02-10 collective_sun/sun_perturbation.tex
remove restriction of zero-mena coefficients in the products of two single- and two-body operators

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v_X$ can be
expanded into its projections $\v v_X^\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v_X = \sum_\Delta \v v_X^\Delta$, which also allows us to
expand
\begin{align}
  \V_1 = \sum_\Delta \V_1^\Delta,
  &&
  \V_1^\Delta \equiv \sum_X \v v_X^\Delta \c \v X,
\end{align}
where each operator $\V_1^\Delta$ generates a state with definite
interaction energy $\Delta$ above that of the permutationally
symmetric manifold $\M_0$.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = - \P_0 \V_1 \E \V_1 \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta}
  \P_0 \V_1^\Delta \P_\Delta \V_1^\Delta \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta} \P_0 \p{\V_1^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PXYP}.  In total, we find
\begin{align}
  H_1^{(2)}
  \EQPS \f1{N\p{N-1}} \sum_{X,Y} \sum_{\Delta\ne0}
  \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
  \p{\col{X}\,\col{Y} - N \col{XY}},
  \label{eq:H_1_2}
\end{align}
where $\EQPS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; $\v v_X^\Delta$ is the projection of
$\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$;
and for any single-spin operator $Z$, we define the collective
operator
\begin{align}
  \col{Z} \equiv \sum_p Z_p.
\end{align}

@@

If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
\end{align}
which implies that $\V_2^\Delta\ket\psi$ is a state of definite energy
that depends only on the coefficients $\m w_O^\Delta$.  Similarly to
the case of single-body perturbations, we can decompose any vector
$\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto the
eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
us to expand
\begin{align}
  \V_2 = \sum_\Delta \V_2^\Delta,
  &&
  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2 \E \V_2 \P_0
  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
Appendix \ref{sec:POQP}.  In the special case that
\begin{align}
  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
\end{align}
where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
the second-order effective Hamiltonian $H_2^{(2)}$ takes the form
\begin{align}
  H_2^{(2)} \simeq_\PS W_4 S_\z^4 - W_2 S_\z^2,
\end{align}
where $\simeq_\PS$ denotes equality up to scalar terms and a
restriction to the permutationally symmetric manifold;
$S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin operator; and
$W_k$ are scalar coefficients.  In order to expand the coefficients
$W_k$, for any ${N \choose 2}$-component vector
$\v{\m m}=\sum_{p<q}m_{pq}\ket{\set{p,q}}$ we define the $N$-component
vector
\begin{align}
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
where we define $m_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $m_{pq}=m_{qp}$ and $m_{pp}=0$.  We then denote the
projection of $\v{\m m}$ onto the $\Delta$-eigenspace of the matrix
$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m m}_\Delta$,
which allows us to concisely write
\begin{align}
  W_4
  \equiv \sum_{\Delta\ne0} \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta}{N!_4},
\end{align}
and
\begin{align}
  W_2
  \equiv \sum_{\Delta\ne0} \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - 2\v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
\end{align}
denotes a falling factorial.

@@

Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
on the permutationally symmetric manifold $\M_0$ by the perturbation
$\V_1$ in \eqref{eq:perturbations} requires us to simplify a product
of the form
\begin{align}
  \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
  \EQPS \diagram{single_body_0} X_1 Y_2
  + \diagram{single_body_1} X_1 Y_1,
  \label{eq:PXYP_start}
\end{align}
where $p,q$ index individual spins; $X,Y$ are single-spin operators;
and the coefficients $v_p,w_q$ satisfy \eqref{eq:cond_1} with
eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
Defining $\v m \equiv \sum_p m_p \ket{p}$ for $m\in\set{v,w}$, we can
simplify
\begin{align}
  \diagram{single_body_1}
  \equiv \sum_p v_p w_p
  = \v v\c\v w,
\end{align}
and
\begin{align}
  \diagram{single_body_0}
  = \diagram{single_body_0_o} - \diagram{single_body_0_x}
  = \diagram{single_body_0_o} - \diagram{single_body_1}
  = - \v v \c\v w,
\end{align}
so
\begin{align}
  \sum_{p,q} v_p w_q X_p Y_q
  \EQPS - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
\end{align}
In order to write this result in terms of collective operators
$\col{Z} \equiv \sum_p Z_p$, we expand
\begin{align}
  \col{X Y} \EQPS N X_1 Y_1,
  &&
  \col{X}\,\col{Y}
  = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
  \EQPS \col{XY} + N\p{N-1} X_1 Y_2,
\end{align}
which implies that
\begin{align}
  \sum_{p,q} v_p w_q X_p Y_q
  \EQPS - \f{\v v \c\v w}{N\p{N-1}}
  \p{\col{X}\,\col{Y} - N \col{XY}}.
\end{align}

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the permutationally symmetric manifold $\M_0$ by the perturbation
$\V_2$ in \eqref{eq:perturbations} requires us to simplify a product
of the form
\begin{multline}
  \P_0 \sp{\sum_{k<\ell} \sum_{p<q}
    v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
  \EQPS \diagram{two_body_0} O_{1,2} Q_{3,4}
  + \diagram{two_body_1} O_{1,2} Q_{1,3}
  + \diagram{two_body_2} O_{1,2} Q_{1,2},
  \label{eq:POQP_start}
\end{multline}
where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
permutationally-symmetric two-spin operators; and the coefficients
$v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta\ne0$, which implies that they are mean-zero.  Defining
\begin{align}
  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
  &&
  \v m \equiv \sum_p m_p \ket{p},
  &&
  m_p \equiv \sum_q m_{pq},
\end{align}
for $m\in\set{v,w}$, we can write
\begin{align}
  \diagram{two_body_2} \equiv \sum_{\p{p,q}\in\C_N\p{2}} v_{pq} w_{pq}
  = \v{\m v} \c\v{\m w},
\end{align}
and simplify
\begin{align}
  \diagram{two_body_1}
  &= \sp{\diagram{two_body_1_o}} - \sp{\diagram{two_body_1_x}} \\
  &= \sp{\diagram{two_body_1_oo} - \diagram{two_body_1_ox}}
  - \sp{2\diagram{two_body_2}} \label{eq:mid_zero} \\
  &= \v v \c \v w - 2 \v{\m v} \c \v{\m w},
\end{align}
where the middle diagram in \eqref{eq:mid_zero} contains an empty sum
for the cross, and is therefore equal to zero.  Similarly neglecting
diagrams with empty sums and additionally using the fact that
$v_{pq},w_{pq}$ are mean-zero, we can work out
\begin{align}
  \diagram{two_body_0}
  &= \f12 \sp{\diagram{two_body_0_o}}
  - \f12 \sp{\diagram{two_body_0_x}} \\
  &= \f12 \sp{\diagram{two_body_0_oo} - \diagram{two_body_0_ox}}
  - \f12 \sp{\diagram{two_body_1}} \\
  &= \f12 \sp{-\diagram{two_body_1_o}}
  - \f12 \p{\v v\c\v w - 2\v{\m v}\c\v{\m w}} \\
  &= -\v v\c\v w + \v{\m v}\c\v{\m w}.
\end{align}
Altogether, we thus have that
\begin{multline}
  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
  \EQPS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
  + \v{\m v} \c \v{\m w}\, O_{1,2} Q_{1,2}.
  \label{eq:POQP}
\end{multline}

@@

Here we simplify a product of three two-body operators projected onto
the permutationally symmetric manifold:
\begin{align}
  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
  &&
  X^2 = 1,
\end{align}
where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
operators that remain after accounting for $X^2=1$, we find that
\begin{align}
  \X \EQPS A_6 X^{\otimes 6} + A_4 X^{\otimes 4}
  + A_2 X^{\otimes 2} + A_0,
  \label{eq:triple_multi}
\end{align}
with
\begin{align}
  A_6 \equiv \diagram{triple_0},
  &&
  A_4 \equiv \diagram{triple_01} + \diagram{triple_1},
  &&
  A_0 \equiv \diagram{triple_0111},
  \label{eq:triple_64}
\end{align}
\begin{align}
  A_2 \equiv \diagram{triple_011} + \diagram{triple_02}
  + \diagram{triple_11} + \diagram{triple_2},
  \label{eq:triple_2}
\end{align}
where a diagram with unlabeled circles denotes a sum over all distinct
label assignments, e.g.
\begin{align}
  \diagram{triple_1} \equiv \diagram{triple_1_uvw},
  \label{eq:triple_uvw}
\end{align}
\begin{align}
  \diagram{triple_011}
  \equiv \diagram{triple_011_uvw}
  + \diagram{triple_011_vwu} + \diagram{triple_011_wuv}.
  \label{eq:triple_uvw_3}
\end{align}
The first diagram in each of $A_6,A_4,A_2$ as written in
\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.

>>>>

We now return to the task of computing the second-order effective
Hamiltonian $H_1^{(2)}$.  Any coefficient vector $\v v_X$ can be
expanded into its projections $\v v_X^\Delta$ onto the eigenspace of
vectors satisfying \eqref{eq:cond_1} with eigenvalue $\Delta$,
i.e.~$\v v_X = \sum_\Delta \v v_X^\Delta$, which also allows us to
expand
\begin{align}
  \V_1 = \sum_\Delta \V_1^\Delta,
  &&
  \V_1^\Delta \equiv \sum_X \v v_X^\Delta \c \v X,
\end{align}
where each operator $\V_1^\Delta$ generates a state with definite
interaction energy $\Delta$ above that of the permutationally
symmetric manifold $\M_0$.  We can therefore simplify
\begin{align}
  H_1^{(2)}
  = - \P_0 \V_1 \E \V_1 \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta}
  \P_0 \V_1^\Delta \P_\Delta \V_1^\Delta \P_0
  = - \sum_{\Delta\ne0} \f1{\Delta} \P_0 \p{\V_1^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_1^\Delta}^2 \P_0$ is worked out in
Appendix \ref{sec:PXYP}.  In total, using the fact that
$\v v_X^\Delta$ is mean-zero for all $\Delta\ne0$, we find
\begin{align}
  H_1^{(2)}
  \EQPS \f1{N\p{N-1}} \sum_{X,Y} \sum_{\Delta\ne0}
  \f{\v v_X^\Delta\c\v v_Y^\Delta}{\Delta}
  \p{\col{X}\,\col{Y} - N \col{XY}},
  &&
  \col{Z} \equiv \sum_p Z_p.
  \label{eq:H_1_2}
\end{align}
where $\EQPS$ denotes equality under a restriction to the fully
symmetric manifold $\M_0$; $\v v_X^\Delta$ is the projection of
$\v v_X$ onto the $\Delta$-eigenspace of the matrix $\m h-\diag\v h$.

@@

If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
\end{align}
which implies that $\V_2^\Delta\ket\psi$ is a state of definite energy
that depends only on the coefficients $\m w_O^\Delta$.  Similarly to
the case of single-body perturbations, we can decompose any vector
$\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto the
eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
us to expand
\begin{align}
  \V_2 = \sum_\Delta \V_2^\Delta,
  &&
  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2 \E \V_2 \P_0
  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
Appendix \ref{sec:POQP}.  In the special case that
\begin{align}
  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
\end{align}
where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
the second-order effective Hamiltonian $H_2^{(2)}$ takes the form
\begin{align}
  H_2^{(2)} \simeq_\PS W_4 S_\z^4 - W_2 S_\z^2,
\end{align}
where $\simeq_\PS$ denotes equality up to scalar terms and a
restriction to the permutationally symmetric manifold;
$S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin operator; and
$W_k$ are scalar coefficients.  In order to expand the coefficients
$W_k$, we define
\begin{align}
  \v{\m w}=\sum_{p<q} w_{pq} \ket{\set{p,q}},
  &&
  \v w \equiv \sum_{p,q} w_{pq} \ket{p},
\end{align}
where $w_{pq}$ is defined for all $p,q$ (as opposed to only $p<q$) by
enforcing $w_{pq}=w_{qp}$ and $w_{pp}=0$.  The vector $\v w$ is the
sum of all columns of the matrix
$\m w\equiv\sum_{p,q}w_{pq}\op{p}{q}$.  We then denote the projection
of $\v{\m w}$ onto the $\Delta$-eigenspace of the matrix
$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m w}_\Delta$,
and define
\begin{align}
  w_{\t{columns}}^{(2)} \equiv \sum_{\Delta\ne0}
  \f{\v w_\Delta\c\v w_\Delta}{\Delta},
  &&
  w_{\t{pairs}}^{(2)} \equiv \sum_{\Delta\ne0}
  \f{\v{\m w}_\Delta\c\v{\m w}_\Delta}{\Delta},
\end{align}
in terms of which
\begin{align}
  W_4 \equiv \f{w_{\t{columns}}^{(2)} - w_{\t{pairs}}^{(2)}}{N!_4},
  &&
  W_2 \equiv \f{w_{\t{columns}}^{(2)} - 2 w_{\t{pairs}}^{(2)}}
  {4N!_2} + \f12 \p{3N-4} W_4,
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
\end{align}
denotes a falling factorial.

@@

Here we simplify the product
\begin{align}
  \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
  \EQPS \diagram{single_body_0} X_1 Y_2
  + \diagram{single_body_1} X_1 Y_1,
  \label{eq:PXYP_start}
\end{align}
where $\P_0$ is a projector onto the permutationally symmetric
manifold $\M_0$; $p,q$ index individual spins; $v_p,w_p$ are scalars;
and $X,Y$ are single-spin operators.  This product appears, for
example, in the calculation of the second-order effective Hamiltonian
$H_1^{(2)}$ induced on the permutationally symmetric manifold $\M_0$
by the single-body perturbation $\V_1$ in \eqref{eq:perturbations}.
Defining
\begin{align}
  \v m \equiv \sum_p m_p \ket{p},
  &&
  \col{m} \equiv \sum_p m_p,
\end{align}
for $m\in\set{v,w}$, we can simplify
\begin{align}
  \diagram{single_body_1}
  \equiv \sum_p v_p w_p
  = \v v\c\v w,
\end{align}
and
\begin{align}
  \diagram{single_body_0}
  = \diagram{single_body_0_o} - \diagram{single_body_0_x}
  = \diagram{single_body_0_oo} - \diagram{single_body_1}
  = \col{u}\,\col{v} - \v v \c\v w,
\end{align}
so
\begin{align}
  \sum_{p,q} v_p w_q X_p Y_q
  \EQPS \col{u}\,\col{v} X_1 Y_2 - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
\end{align}
In order to write this result in terms of collective operators
$\col{Z} \equiv \sum_p Z_p$, we expand
\begin{align}
  \col{X Y} \EQPS N X_1 Y_1,
  &&
  \col{X}\,\col{Y} \EQPS \col{XY} + N\p{N-1} X_1 Y_2,
\end{align}
which implies that
\begin{align}
  \sum_{p,q} v_p w_q X_p Y_q
  \EQPS \f{\col{u}\,\col{v}}{N\p{N-1}}
  \p{\col{X}\,\col{Y} - \col{XY}}
  - \f{\v v \c\v w}{N\p{N-1}}
  \p{\col{X}\,\col{Y} - N\col{XY}}.
\end{align}

@@

Here we simplify the product
\begin{multline}
  \P_0 \sp{\sum_{k<\ell} \sum_{p<q}
    v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
  \EQPS \diagram{two_body_0} O_{1,2} Q_{3,4}
  + \diagram{two_body_1} O_{1,2} Q_{1,3}
  + \diagram{two_body_2} O_{1,2} Q_{1,2},
  \label{eq:POQP_start}
\end{multline}
where $\P_0$ is a projector onto the permutationally symmetric
manifold $\M_0$; $k,\ell,p,q$ index individual spins;
$v_{k\ell},w_{pq}$ are scalar coefficients with $m_{pq}=m_{qp}$ and
$m_{pp}=0$ for each of $m\in\set{v,w}$; and $O,Q$ are
permutationally-symmetric two-spin operators.  This product appears,
for example, in the calculation of the second-order effective
Hamiltonian $H_2^{(2)}$ induced on the permutationally symmetric
manifold $\M_0$ by the two-body perturbation $\V_2$ in
\eqref{eq:perturbations} in \eqref{eq:perturbations}.  Defining
\begin{align}
  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
  &&
  \v m \equiv \sum_p m_p \ket{p},
  &&
  m_p \equiv \sum_q m_{pq},
  &&
  \col{m} \equiv \sum_{p<q} m_{pq},
\end{align}
for $m\in\set{v,w}$, we can write
\begin{align}
  \diagram{two_body_2} \equiv \sum_{\p{p,q}\in\C_N\p{2}} v_{pq} w_{pq}
  = \v{\m v} \c\v{\m w},
\end{align}
and simplify
\begin{align}
  \diagram{two_body_1}
  &= \sp{\diagram{two_body_1_o}} - \sp{\diagram{two_body_1_x}} \\
  &= \sp{\diagram{two_body_1_oo} - \diagram{two_body_1_ox}}
  - \sp{2\diagram{two_body_2}} \label{eq:mid_zero} \\
  &= \v v \c \v w - 2 \v{\m v} \c \v{\m w},
\end{align}
where the middle diagram in \eqref{eq:mid_zero} contains an empty sum
for the cross, and is therefore equal to zero.  Finally, with a bit of
more work (or a computer) we can work out that
\begin{align}
  \diagram{two_body_0}
  = \diagram{two_body_0_oooo} - \diagram{two_body_1_ooo}
  + \diagram{two_body_2_oo}
  = \col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}.
\end{align}
Altogether,
\begin{multline}
  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
  \EQPS \p{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}
  O_{1,2} Q_{3,4}
  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
  + \v{\m v} \c \v{\m w}\, O_{1,2} Q_{1,2}.
  \label{eq:POQP}
\end{multline}
In the special case that $O=Q=X\otimes X$ with $X^2=1$, we can further
simplify
\begin{align}
  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} X_k X_\ell X_p X_q
  \simeq_\PS
  \p{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}} X^{\otimes 4}
  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} X^{\otimes 2},
\end{align}
where $\simeq_\PS$ denotes equality up to scalar terms and a
restriction to the permutationally symmetric manifold.  In terms of
collective operators $\col{X} \equiv \sum_p X_p$, we then have
\begin{multline}
  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} X_k X_\ell X_p X_q \\
  \simeq_\PS
  \f{\col{v}\,\col{w} -\v v\c\v w + \v{\m v}\c\v{\m w}}{N!_4}
  \sp{\col{X}^4 - 2\p{3N-4} \col{X}^2}
  + \f{\p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}}}{N!_2}\, \col{X}^2,
\end{multline}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
\end{align}
denotes a falling factorial.

@@

Here we simplify the product
\begin{align}
  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
  &&
  X^2 = 1,
\end{align}
where $\P_0$ is a projector onto the permutationally symmetric
manifold $\M_0$; $k,\ell,p,q,r,s$ index individual spins; $X$ is a
single-spin operator; and $u_{k\ell},v_{pq},w_{rs}$ are scalar
coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
operators that remain after accounting for $X^2=1$, we find that
\begin{align}
  \X \EQPS A_6 X^{\otimes 6} + A_4 X^{\otimes 4}
  + A_2 X^{\otimes 2} + A_0,
  \label{eq:triple_multi}
\end{align}
with
\begin{align}
  A_6 \equiv \diagram{triple_0},
  &&
  A_4 \equiv \diagram{triple_01} + \diagram{triple_1},
  &&
  A_0 \equiv \diagram{triple_0111},
  \label{eq:triple_64}
\end{align}
\begin{align}
  A_2 \equiv \diagram{triple_011} + \diagram{triple_02}
  + \diagram{triple_11} + \diagram{triple_2},
  \label{eq:triple_2}
\end{align}
where a diagram with unlabeled circles denotes a sum over all distinct
label assignments, e.g.
\begin{align}
  \diagram{triple_1} \equiv \diagram{triple_1_uvw},
  \label{eq:triple_uvw}
\end{align}
\begin{align}
  \diagram{triple_011}
  \equiv \diagram{triple_011_uvw}
  + \diagram{triple_011_vwu} + \diagram{triple_011_wuv}.
  \label{eq:triple_uvw_3}
\end{align}
The first diagram in each of $A_6,A_4,A_2$ as written in
\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.
